---
title: "DACSS695_SR_Final Project"
format: html
editor: visual
---

## 695SR Final Project Survey Analysis

### Loading the Data

```{r setup, include=FALSE}
library(haven)
library(ggplot2)
library(dplyr)
library(tidyr)
library(knitr)
library(tidyverse)
library(labelled)
library(descr)
library(srvyr)
library(survey)
library(weights)
library(anesrake)
library(data.table)
library(stargazer)
library(rstatix)
library(skimr)
library(janitor)
library(pollster)
library(corrplot)
library(gridExtra)
library(kableExtra)
```

```{r}
url <- "https://github.com/drCES/raw_survey_data_dacss695/raw/main/data_clean.dta"
data <- read_dta(url)# Read the Stata file into R
data <- as.data.frame(data) #for anesrake to work, data must be in data.frame
```

### Reviewing Unweighted Sample Demographic Profiles

```{r}
names(data)
```

```{r}
freq(data$age)
val_labels(data$age)
freq(data$gender)
val_labels(data$gender)
freq(data$ethnicity)
val_labels(data$ethnicity)
freq(data$hispanic)
val_labels(data$hispanic)
freq(data$education)
val_labels(data$education)
```

### Recoding existing data to match population target values

```{r}
#class example, recoding existing data to match pop. target values
data <- data %>%
  mutate(college = if_else(educ<=5, 2, 1),
         hispanic2 = if_else(hispanic>1, 1, 2),
         age_group = case_when(
           age >= 18 & age <= 25 ~ 1,
           age >= 26 & age <= 34 ~ 2,
           age >= 35 & age <= 54 ~ 3,
           age >= 55 & age <= 64 ~ 4, 
           age > 64  ~ 5),
         race5 = case_when(
           ethnicity == 1  ~ 1,
           ethnicity == 2  ~ 2,
           ethnicity >= 4 & ethnicity <= 10 ~ 3,
           ethnicity == 3 ~ 4, 
           ethnicity >= 11 & ethnicity <= 15 ~ 5)
         )
```

```{r}
freq(data$gender)
freq(data$age_group)
freq(data$race5)
freq(data$hispanic2)
freq(data$college)
```

```{r}
#doing value labels where i need them for toplines and crosstabs
data <- data %>% 
  mutate(age_group = labelled(age_group, labels = c(
    "18-25" = 1, 
    "26-34" = 2, 
    "35-54"= 3, 
    "55-64" = 4, 
    "65+" = 5
    )))

val_labels(data$age_group) 
```

### Creating the weights

```{r}
#Creating Survey Weight Vectors
#gender, age_group, race5, hispanic2, college
age_group <- c(0.113, 0.160, 0.334, 0.168, 0.225) 
sum(age_group) 
gender <-c(0.483, 0.517) 
sum(gender) 
race5 <- c(0.641, 0.120, 0.061, 0.013, 0.165)
sum(race5)
hispanic2 <- c(0.168, 0.832)
sum(hispanic2)
college <- c(0.377, 0.623)
sum(college)
```

### Calculating Your Post-Stratification Weights

```{r}
#gender, age_group, race5, hispanic2, college
#Step 1: Save the target list 
targets <- list(age_group, gender, race5, hispanic2, college)
# remember, these names will have to match
names(targets) <- c("age_group", "gender", "race5", "hispanic2","college")
```

```{r}
#Step 2 - Calculate the Rake Weight
#set.seed(1967) #Set the seed for replication  
myweights <- anesrake(targets, data, 
  caseid = data$response_id, cap = 8, type = "pctlim", pctlim=.05)
```

```{r}
#Step 3 - Save the Rake Weight at the end of your sample data
data$full_weight  <- unlist(myweights[1])

#Step 4: Reviewing the newly created weights
summary(data$full_weight) #Displays summary of the weight size to see range
```

### Creating topline reports and cross tabulations

```{r}
#frequency tables
freq(data$gutierrez_suarez_ai, plot=F)
freq(data$ai_ad_2, plot=F)
freq(data$gutierrez_clear, plot=F)
```

My 3 topline reports

```{r gutierrez_suarez_ai}
topline(data, variable = gutierrez_suarez_ai, 
        weight = full_weight, pct = FALSE, cum_pct=FALSE) %>%  
  kable(digits=0, 
        col.names = c('Knowledge', 'N', 
                      'Percent Selected'),
        escape=FALSE, align=('lcc')) %>%  
  kable_styling(bootstrap_options = "responsive", full_width = F) %>%
  kableExtra::footnote(alphabet = ("Full Sample MOE +/-6%")) %>%
  kableExtra::add_header_above(c("Respondent Knowledge on Artificial Intelligence" = 3))
```

```{r ai_ad_2}
topline(data, variable = ai_ad_2, 
        weight = full_weight, pct = FALSE, cum_pct=FALSE) %>%  
  kable(digits=0, 
        col.names = c('Agree or Disagree', 'N', 
                      'Percent Selected'),
        escape=FALSE, align=('lcc')) %>%  
  kable_styling(bootstrap_options = "responsive", full_width = F) %>%
  kableExtra::footnote(alphabet = ("Full Sample MOE +/-6%")) %>%
  kableExtra::add_header_above(c("Statement: The integration of AI into our daily life is a good thing overall." = 3))
```

```{r gutierrez_clear}
topline(data, variable = gutierrez_clear, 
        weight = full_weight, pct = FALSE, cum_pct=FALSE, remove = c("(Missing)")) %>%  
  kable(digits=0, 
        col.names = c('Levels of Importance', 'N', 
                      'Percent Selected'),
        escape=FALSE, align=('lcc')) %>%  
  kable_styling(bootstrap_options = "responsive", full_width = F) %>%
  kableExtra::footnote(alphabet = ("Full Sample MOE +/-6%")) %>%
  kableExtra::add_header_above(c("Respondent Attitudes Toward AI Providing Clear Explanations" = 3))
```

My interesting crosstab(s) to look at how different subgroups in my sample answered my key question of interest.

```{r}
#looking at a 2-way crosstable using gutierrez_suarez_ai (ai knowledge) as the key outcome variable of interest by the variable college
crosstab(df = data, x = gutierrez_suarez_ai, 
         y = college, weight = full_weight, 
         pct_type = "col") %>%
  kable(digits=0,  
        col.names = c('Knowledge','No college education', 'College education'), align=('lccccr')) %>%
  kable_styling(bootstrap_options = "responsive", full_width = F, position = "left") %>% 
  kableExtra::footnote(alphabet = ("Full Sample MOE +/-6%; Values Represent Percentage per Knowledge Level")) %>%
  kableExtra::add_header_above(c("Respondent Knowledge on Artifical Intelligence by College Education" = 3))

```

```{r}
#looking at a 2-way crosstable using gutierrez_suarez_ai (ai knowledge) as the key outcome variable of interest by the variable age_group
crosstab(df = data, x = gutierrez_suarez_ai, 
         y = age_group, weight = full_weight, 
         pct_type = "col") %>%
  kable(digits=0,  
        col.names = c('Knowledge', '18-25', '26-34', '35-54', '55-64', '65+'), align=('lccccr')) %>%
  kable_styling(bootstrap_options = "responsive", full_width = F, position = "left") %>% 
  kableExtra::footnote(alphabet = ("Full Sample MOE +/-6%; Values Represent Percentage per Knowledge Level")) %>%
  kableExtra::add_header_above(c("Respondent Knowledge on Artifical Intelligence by Age Group" = 6))
```

```{r}
#recoding political_party variable before crosstabing
data <- data %>%
  mutate(political_party3 = case_when(
           political_party == 1  ~ 1,
           political_party == 2 ~ 2,
           political_party == 3 ~ 3)
         )

data <- data %>% 
  mutate(political_party3 = labelled(political_party3, labels = c(
    "Democrat" = 1, 
    "Republican" = 2, 
    "Independent"= 3
    )))

freq(data$political_party3, plot=F)
val_labels(data$political_party3) 
```

```{r}
#looking at a 2-way crosstable using ai_ad_2 as the key outcome variable of interest by political_party
#Statement: The integration of AI into our daily life is a good thing overall
crosstab(df = data, x = ai_ad_2, 
         y = political_party3, weight = full_weight, 
         pct_type = "col") %>%
  kable(digits=0,  
        col.names = c('Response', 'Democrat', 'Republican', 'Independent'), align=('lccccr')) %>%
  kable_styling(bootstrap_options = "responsive", full_width = F, position = "left") %>% 
  kableExtra::footnote(alphabet = ("Full Sample MOE +/-6%; Values Represent Percentage per Response")) %>%
  kableExtra::add_header_above(c("Respondent Attitudes by Political Party on the Statement: The integration of AI into our daily life is a good thing overall." = 4))
```

```{r}
#looking at a 2-way crosstable using gutierrez_clear (ai clear statements) as the key outcome variable of interest by the variable college
crosstab(df = data, x = gutierrez_clear, 
         y = college, weight = full_weight, 
         pct_type = "col") %>%
  kable(digits=0,  
        col.names = c('Levels of Importance','No college education', 'College education'), align=('lccccr')) %>%
  kable_styling(bootstrap_options = "responsive", full_width = F, position = "left") %>% 
  kableExtra::footnote(alphabet = ("Full Sample MOE +/-6%; Values Represent Percentage per Selected Level")) %>%
  kableExtra::add_header_above(c("How important is it for you that AI systems provide clear explanations for their responses or decisions, if at all?" = 3))
```
